---
title: "01_acoustic_index_function"
output: html_document
date: "2026-02-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tuneR)
library(seewave)
library(soundecology)
library(dplyr)
library(lubridate)
```



## raw data (audio files) accessed through microsoft azure storage - not stored directly within repository


# Acoustic Indices

```{r}

allindicesfunc<-function(weekstarting,date,sites){
  #create big function to calculate indices for all named sites on a given date
  
  for(x in sites){
    #go through site A, then site B etc
    
    weekstartingsite<-paste(weekstarting,x,sep="/")#make weekstarting/A, weekstarting/B etc
    
    #files with week starting and site
    filepart1<-paste("Z:/AudioMoth_data",weekstartingsite,sep="/")
    
    #pathway with week starting and site
    pathway1 <- paste("Z:/AudioMoth_data",weekstartingsite,sep="/")
    
 allvaluesdf  <- NULL;
    for(n in 0:2){
      #go through 0,1,2 for each site
  
  ##make file names to read in
      #files with exact site
filepart2<-paste(filepart1,n,sep="") #make A0, A1 etc

#files with exact site and date
filenames<-paste(filepart2,date,sep="/")
#list of files with exact site and date
files<-list.files(filenames)

  
  #make file pathway
#location of the listed files
#pathway with exact site
pathway2<-paste(pathway1,n,sep="") #make A0, A1 etc
#pathway with exact site and date
pathways<-paste(pathway2,date,sep="/")
  

#function to calculate indices for each file
indicesfunc<-function(file,filepath) { 

 # file names to read in - pathway followed by name of individual file
  name <- paste(filepath, file,  sep="/")
  #read in the wav files
  wav <- readWave(name)

  #Bioacoustic Index
 BIO<- bioacoustic_index(wav, min_freq = 2000, max_freq = 10000)$left_area
  
  #Acoustic Complexity Index
 ACI<- ACI(wav, flim=c(2,10))
  
  #Normalised Difference Soundscape Index
 NDSI<- NDSI(soundscapespec(wav, plot=FALSE))
  
#Create Data Frame Including File Name and Index Values
  valuedf<-data.frame(file,BIO,ACI,NDSI,name)
  

}#end of indicesfunc

#run the above function for all files in the folders for 0,1,2 for a given letter and bind data frames together
#make data frame of indices for all files in the folder
 valuesdf <- do.call(rbind.data.frame,
             lapply(files,indicesfunc,pathways))
 #compile outputs for locations 0,1,2 for each file pathway
 allvaluesdf <- rbind(allvaluesdf, valuesdf)
allvaluesdf 

}#end of second for loop - function has run for 0,1,2 for a given site

allvaluesdf
  
#make site variable in data frame by extracting site name from file name
allvaluesdf$site<-allvaluesdf$name
#characters 40 and 41 of file name
 allvaluesdf$site<-substr(allvaluesdf$site,40,41)
 
  #make datetime variable by extracting datetime from file name
allvaluesdf$datetime<-allvaluesdf$file
allvaluesdf$datetime= substr(allvaluesdf$datetime,1,nchar(allvaluesdf$datetime)-4)
allvaluesdf$datetime<-ymd_hms(allvaluesdf$datetime)

#make time variable
allvaluesdf$time<-allvaluesdf$datetime
allvaluesdf$time<-format(strptime(allvaluesdf$time, "%Y-%m-%d %H:%M:%S"), "%H:%M:%S")

  #make output path
#paste site name and date together to get unique name for a given day of data at a given site
outputname1<-paste(x,date,sep="")
#add .csv to the end of the name so the output can be written as a csv file
outputname<-paste(outputname1,".csv",sep="")

#pathway for where to put the csv files with indices in them
outputfolder<-paste("kelvin_corridor_acoustic/outputs/acoustic-indices",weekstarting,sep="/")

outputlocation<-paste(outputfolder,outputname,sep="/")




#write csv file with indices for a given site
write.csv(allvaluesdf,outputlocation)



    }#end of first for loop
    #at this point the above function has run for each site separately, generating separate csv files for each site

}



##run function for test data 
##Week_starting_230417
##A & B
##20230409
allindicesfunc("Week_starting_230417","20230409",c("A","B"))

###running this function writes csv files for A0,A1,A2,B0,B1,B2 day 20230409 with all 3 acoustic indices

```



For the indices function to work, the sites listed for a given date must be in the same week_starting folder. If they are in different week_starting folders due to different SD card swap dates across the sites, the sites in different folders must be run separately.

It is probably necessary to add an empty folder if there is no data from a particular location so the function doesn't error - eg if on a given date there is data from all sites except D2, you should probably add folders as follows: week_starting_date/D2/date




run this function for all days of data to process
