---
title: "03_acoustic_index_ground_truthing"
output: html_document
date: "2026-02-05"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = 'C:/Users/2298350G/Documents/My Documents/GALLANT project/kelvin_corridor_acoustic')
library(tuneR)
library(seewave)
library(soundecology)
library(dplyr)
library(lubridate)
library(ggplot2)
library(glmmTMB)
library(pscl)
library(lmtest)
```


#calculate indices for each validation subsample file

```{r}
#function to calculate indices for each file
testindicesfunc<-function(file,filepath) { 

 # file names to read in - pathway followed by name of individual file
  name <- paste(filepath, file,  sep="/")
  #read in the wav files
  wav <- readWave(name)

  #Bioacoustic Index
 BIO<- bioacoustic_index(wav, min_freq = 2000, max_freq = 10000)$left_area
  
  #Acoustic Complexity Index
 ACI<- ACI(wav, flim=c(2,10))
  
  #Normalised Difference Soundscape Index
 NDSI<- NDSI(soundscapespec(wav, plot=FALSE))
  
#Create Data Frame Including File Name and Index Values
  valuedf<-data.frame(file,BIO,ACI,NDSI,name)
  

}#end of indicesfunc

#list of files with exact site and date
testfiles<-list.files("kelvin_corridor_acoustic/validation_subset")

testpathway<-"kelvin_corridor_acoustic/validation_subset"

#run the above function for all files in the folders for 0,1,2 for a given letter and bind data frames together
#make data frame of indices for all files in the folder
 allvaluesdf <- do.call(rbind.data.frame,
             lapply(testfiles,testindicesfunc,testpathway))


#end of second for loop - function has run for 0,1,2 for a given site

allvaluesdf
  

#convert NDSI to values between 0 and 1 rather than -1 and 1
allvaluesdf$NDSIoriginal<-allvaluesdf$NDSI
allvaluesdf$NDSI<-(allvaluesdf$NDSI+1)/2

#rescale BIO
allvaluesdf$BIOrescale <- (allvaluesdf$BIO - min(allvaluesdf$BIO)) / (max(allvaluesdf$BIO) - min(allvaluesdf$BIO))

#rescale ACI
allvaluesdf$ACIrescale <- (allvaluesdf$ACI - min(allvaluesdf$ACI)) / (max(allvaluesdf$ACI) - min(allvaluesdf$ACI))

#make mean of all scaled acoustic indices
allvaluesdf$meanindex<-(allvaluesdf$ACIrescale+allvaluesdf$BIOrescale+allvaluesdf$NDSI)/3

#write csv file with indices
write.csv(allvaluesdf,"kelvin_corridor_acoustic/outputs/ground_truthing/validation_indices.csv")


```




# verification analysis


## read in manual classifications of each recording

```{r}
testdata<-read.csv("data/validation_summary.csv", header=T, sep="," )
testdata$mean2<-(testdata$NDSI+testdata$BIOrescale)/2
```


try correlation tests of each acoustic index with biophony_diversity, biophony_duration, biophony_intensity_number, anthrophony_cat_num, anthrophony_duration, geophony_cat_num, geophony_duration


## test normality of variables

```{r}
#acoustic indices
hist(testdata$NDSI,breaks = seq(min(testdata$NDSI), max(testdata$NDSI),length.out=20))
shapiro.test(testdata$NDSI)#not normally distributed

hist(testdata$ACIrescale,breaks = seq(min(testdata$ACIrescale), max(testdata$ACIrescale),length.out=20))
shapiro.test(testdata$ACIrescale)#not normally distributed

hist(testdata$BIOrescale,breaks = seq(min(testdata$BIOrescale), max(testdata$BIOrescale),length.out=20))
shapiro.test(testdata$BIOrescale)#not normally distributed

hist(testdata$meanindex,breaks = seq(min(testdata$meanindex), max(testdata$meanindex),length.out=20))
shapiro.test(testdata$meanindex)#not normally distributed
```


## kendall's correlation tests


## use p.adjust function for each p value (Bonferroni correction for n=18)


### NDSI

```{r}
cor.test(testdata$NDSI,testdata$biophony_diversity,method="kendall")#insignificant
#p.adjust(p, n=18)

cor.test(testdata$NDSI,testdata$biophony_duration,method="kendall")#significant

cor.test(testdata$NDSI,testdata$biophony_intensity_number,method="kendall")#significant

cor.test(testdata$NDSI,testdata$anthrophony_cat_num,method="kendall")#significant

cor.test(testdata$NDSI,testdata$anthrophony_duration,method="kendall")#significant

cor.test(testdata$NDSI,testdata$geophony_cat_num,method="kendall")#insignificant

cor.test(testdata$NDSI,testdata$geophony_duration,method="kendall")#insignificant

#NDSI shows significant correlation with biophony duration & intensity and anthrophony diversity and duration
```



### ACI

```{r}
cor.test(testdata$ACIrescale,testdata$biophony_diversity,method="kendall")#insignificant

cor.test(testdata$ACIrescale,testdata$biophony_duration,method="kendall")#insignificant

cor.test(testdata$ACIrescale,testdata$biophony_intensity_number,method="kendall")#insignificant

cor.test(testdata$ACIrescale,testdata$anthrophony_cat_num,method="kendall")#insignificant

cor.test(testdata$ACIrescale,testdata$anthrophony_duration,method="kendall")#insignificant

cor.test(testdata$ACIrescale,testdata$geophony_cat_num,method="kendall")#significant - positive relationship

cor.test(testdata$ACIrescale,testdata$geophony_duration,method="kendall")#significant - positive relationship

#ACI shows significant positive correlation with geophony diversity and duration

```


### BIO

```{r}
cor.test(testdata$BIOrescale,testdata$biophony_diversity,method="kendall")#significant
 
cor.test(testdata$BIOrescale,testdata$biophony_duration,method="kendall")#significant

cor.test(testdata$BIOrescale,testdata$biophony_intensity_number,method="kendall")#significant

cor.test(testdata$BIOrescale,testdata$anthrophony_cat_num,method="kendall")#insignificant

cor.test(testdata$BIOrescale,testdata$anthrophony_duration,method="kendall")#insignificant

cor.test(testdata$BIOrescale,testdata$geophony_cat_num,method="kendall")#insignificant 

cor.test(testdata$BIOrescale,testdata$geophony_duration,method="kendall")#insignificant

#BIO shows significant relationship with biophony diversity, duration and intensity
```


### mean index

```{r}
cor.test(testdata$meanindex,testdata$biophony_diversity,method="kendall")#significant

cor.test(testdata$meanindex,testdata$biophony_duration,method="kendall")#significant

cor.test(testdata$meanindex,testdata$biophony_intensity_number,method="kendall")#significant

cor.test(testdata$meanindex,testdata$anthrophony_cat_num,method="kendall")#nearly significant

cor.test(testdata$meanindex,testdata$anthrophony_duration,method="kendall")#significant

cor.test(testdata$meanindex,testdata$geophony_cat_num,method="kendall")#insignificant 

cor.test(testdata$meanindex,testdata$geophony_duration,method="kendall")#insignificant


#mean acoustic index shows significant correlation with biophony diversity, duration and intensity, and anthrophony duration, and marginally significant correlation with anthrophony diversity
```


### mean of NDSI and BIO

```{r}
cor.test(testdata$mean2,testdata$biophony_diversity,method="kendall")#significant

cor.test(testdata$mean2,testdata$biophony_duration,method="kendall")#significant

cor.test(testdata$mean2,testdata$biophony_intensity_number,method="kendall")#significant

cor.test(testdata$mean2,testdata$anthrophony_cat_num,method="kendall")#insignificant

cor.test(testdata$mean2,testdata$anthrophony_duration,method="kendall")#significant

cor.test(testdata$mean2,testdata$geophony_cat_num,method="kendall")#insignificant 

cor.test(testdata$mean2,testdata$geophony_duration,method="kendall")#insignificant

```



bioacoustic index shows strongest correlation with biophony metrics



plots

plot NDSI against biophony duration and anthrophony duration

plot BIO against biophony duration and diversity


```{r}
NDSIplot1 <- ggscatter(testdata, x = "biophony_duration", y = "NDSI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Biophony Duration (s)", ylab = "NDSI")

NDSIplot2 <- ggscatter(testdata, x = "anthrophony_duration", y = "NDSI", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Anthrophony Duration (s)", ylab = "NDSI")

BIOplot1 <- ggscatter(testdata, x = "biophony_duration", y = "BIOrescale", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Biophony Duration (s)", ylab = "BIO")

BIOplot2 <- ggscatter(testdata, x = "biophony_diversity", y = "BIOrescale", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "kendall",
          xlab = "Biophony Diversity", ylab = "BIO")



top_row = ggarrange(NDSIplot1, NDSIplot2, ncol = 2, labels = c("A", "B"))
bottom_row = ggarrange(BIOplot1, BIOplot2, ncol = 2, labels = c("C", "D"))
final_plot = ggarrange(top_row, bottom_row, ncol = 1)
final_plot

ggsave(filename = "outputs/figures/index_validation.png",
       plot = final_plot,
       device = "png",
       width = 20,
       height = 15,
       units = "cm",
       dpi = 300,
       bg = "white")

```


