---
title: "04_environmental_variables"
output: html_document
date: "2026-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(ggplot2)
library(data.table)
library(lme4)
library(lmtest)
library(suntools)
library(betareg)
library(circular)
library(data.table)
library(glmmTMB)
library(betareg)
library(mgcv)
library(circular)
```




# read in data

```{r}
indices<- fread("kelvin_corridor_acoustic/outputs/full-data/allindices.csv", header=T, sep="," )

indicesdf<-subset(indices,select=-c(V1,X,name,file))
#replace NA values in time variable with 00:00:00
indicesdf$time[is.na(indicesdf$time)] <- "00:00:00"
#make time variable as seconds after midnight
indicesdf$ntime<-as.numeric(hms(indicesdf$time))

#make date variable
indicesdf$date<-substring(indicesdf$datetime,1,10)
indicesdf$date<-ymd(indicesdf$date)
#make julian date variable
indicesdf$julian <- strftime(indicesdf$date, format = "%j")
indicesdf$julian<-as.numeric(indicesdf$julian)

#convert NDSI to values between 0 and 1 rather than -1 and 1
indicesdf$NDSIoriginal<-indicesdf$NDSI
indicesdf$NDSI<-(indicesdf$NDSI+1)/2

indicesdf$site<-as.factor(indicesdf$site)

```



## read in GIS site data

```{r}
sitedata<- fread("kelvin_corridor_acoustic/data/environmental_variables/GIS-site-data.csv", header=T, sep="," )
sitedata$site<-as.factor(sitedata$site)

indicesdf<-merge(indicesdf,sitedata,by="site")

#remove rows with NA values
indicesdf<-indicesdf[complete.cases(indicesdf), ]


indicesdf$timemin<-(indicesdf$ntime)/60
indicesdf$timehr<-(indicesdf$timemin)/60
indicesdf$ntime2<-(indicesdf$ntime)^2
indicesdf$timehr2<-(indicesdf$timehr)^2

#log transform kelvindist
indicesdf$logKelvindist<-log(indicesdf$Kelvindist)

indicesdf$Clydedistkm<-indicesdf$Clydedist/1000

```



## land cover data

```{r}
lcdata<- read.csv("kelvin_corridor_acoustic/data/environmental_variables/land_cover_2.csv", header=T, sep="," )

lc2<-lcdata[,2:13]

lc2 <- asin(sqrt(lc2[,1:12]/100))

lc2pca <- prcomp(lc2, scale=T)
summary(lc2pca)

plot(lc2pca, type="lines")
lc2pca$rotation
lc2pcload<-as.data.frame(round(lc2pca$rotation*100))

lcPC <- data.frame(lc2pca$x)
plot(lcPC$PC1, lcPC$PC2)

lc3<-cbind(lcdata$site,lcPC)
lc3$site<-lc3$`lcdata$site`
lc4<-subset(lc3,select=c(site,PC1,PC2,PC3,PC4))
lc4$site<-as.factor(lc4$site)
indicesdf<-merge(indicesdf,lc4,by="site")

lcdata$site<-as.factor(lcdata$site)
indicesdf<-merge(indicesdf,lcdata,by="site")
```










## Met Office data from Bishopton weather station

```{r}
t1<-read.csv("temp/Bishopton March-April.csv", header=T, sep="," ) 
t2<-read.csv("temp/Bishopton April-May.csv", header=T, sep="," ) 
t3<-read.csv("temp/Bishopton May-June.csv", header=T, sep="," ) 
t4<-read.csv("temp/Bishopton June-July.csv", header=T, sep="," ) 
t5<-read.csv("temp/Bishopton July-August.csv", header=T, sep="," ) 
t6<-read.csv("temp/Bishopton August-September.csv", header=T, sep="," ) 

allt<-rbind(t1,t2,t3,t4,t5,t6)
allt$date<-substr(allt$Report.Date...Time,1,10)
allt$time<-substr(allt$Report.Date...Time,12,16)
allt$timet<-allt$time
allt[allt == "" | allt == " "] <- NA  # Replace blank & space by NA
allt$timet[is.na(allt$timet)] <- "00:00"
allt$datetime<-paste(allt$date,allt$timet,sep="_")
allt$datetime<-dmy_hm(allt$datetime)

metdata<-subset(allt,select=c(datetime,Air.Temperature,Relative.Humidity))
```



## read in temperature logger data

```{r}
tempdata<- fread("data/environmental_variables/Kestrel_data_cleaned.csv", header=T, sep="," )


tempdata2<-subset(tempdata,select=c(DateTime,ID,Temp,Rel.Humidity))
tempdata2$datetime<-tempdata2$DateTime

alltemps<-merge(tempdata2,metdata,by="datetime")

alltemps$site<-as.factor(alltemps$ID)

min(alltemps$datetime)
indicesdft<-merge(indicesdf,metdata,by="datetime",all=T)
```



## analysis

```{r}
tmod1<-lm(Temp~Air.Temperature+site*Air.Temperature*site,data=alltemps)
summary(tmod1)
hist(resid(tmod1))
plot(tmod1)

tmod2<-glmmTMB(Temp~Air.Temperature+site*Air.Temperature*site,family=Gamma(link="log"),data=alltemps,REML=FALSE)
summary(tmod2)
hist(resid(tmod2))
tmod2resid <- simulateResiduals(fittedModel = tmod2,re.form=NA)
plotQQunif(tmod2resid)
plotResiduals(tmod2resid)

AIC(tmod1,tmod2)


hmod1<-lm(Rel.Humidity~Relative.Humidity+site*Relative.Humidity*site,data=alltemps)
summary(hmod1)
hist(resid(hmod1))
plot(hmod1)
```



## predict temperatures for full indices data set using linear model

```{r}
temppreds<-predict(tmod1,newdata=indicesdft, se.fit=TRUE,type="response")
temppreds<-as.data.frame(temppreds)

indicesdft$tpred<-temppreds$fit
#add variables for standard errors
indicesdft$minse<-temppreds$fit-(temppreds$se.fit)
indicesdft$maxse<-temppreds$fit+(temppreds$se.fit)



#predict humidity values
hpreds<-predict(hmod1,newdata=indicesdft, se.fit=TRUE,type="response")
hpreds<-as.data.frame(hpreds)

indicesdft$hpred<-hpreds$fit
```



## check correlation between temperature from met office data and predicted temperature

```{r}
cor.test(indicesdft$Air.Temperature,indicesdft$tpred)
#correlation between humidity predicted and observed from met office
cor.test(indicesdft$Relative.Humidity,indicesdft$hpred)

#add temperature data from temperature loggers
indicesdf2temph<-subset(indicesdf2,select=c(sitedatetime,Temp,Rel.Humidity))
indicesdft$sitedatetime<-paste(indicesdft$site,indicesdft$datetime,sep="-")

indicesdftsub=merge(indicesdf2temph,indicesdft,by="sitedatetime")

#correlation between observed from temperature loggers and predicted
cor.test(indicesdftsub$Temp,indicesdftsub$tpred)
cor.test(indicesdftsub$Rel.Humidity,indicesdftsub$hpred)
```


## make data set of indices without temperature logger data

```{r}
indicesdfmin<-filter(indicesdf,!datetime%in%c(indicesdf2$datetime))
indicesdfmin$sitedatetime<-paste(indicesdfmin$site,indicesdfmin$datetime,sep="-")

indicesdft$Temp<-indicesdft$tpred
indicesdft$Rel.Humidity<-indicesdft$hpred
indicesdft2<-subset(indicesdft,select=c(sitedatetime,Temp,Rel.Humidity))

indicesdfmin2<-merge(indicesdfmin,indicesdft2,by="sitedatetime")
indicesdfmin2$pred<-"Y"

indicesdf2


#data set with predicted temperature
#remove datetimehour
indicesdfmin2<-subset(indicesdfmin2,select=-c(datetimehour))


#data set with temperature logger data
#add pred variable
indicesdf2$pred<-"N"
#remove Station.Pressure and Dew.Point
indicesdf2<-subset(indicesdf2,select=-c(Station.Pressure,Dew.Point))
#remove datetimehour and Heat.Index and Density.Altitude
indicesdf2<-subset(indicesdf2,select=-c(Heat.Index,Density.Altitude))
indicesdf2<-subset(indicesdf2,select=-c(datetimehour))


allindices<-rbind(indicesdfmin2,indicesdf2)


## write full data as csv
write.csv(allindices,"kelvin_corridor_acoustic/outputs/full-data/allindices2.csv")
```








# compile all data


## read in data with simulated temperature and humidity values

```{r}
allindices2<-fread("outputs/full-data/allindices2.csv", header=T, sep=",")
str(allindices2)

##add in new clyde distance (river distance)
sitedata<- fread("data/environmental_variables/all_site_data.csv", header=T, sep="," )
sitedata$site<-as.factor(sitedata$site)
sitedataclyde<-subset(sitedata,select=c(site,newClydedist0,newClydedist2))
allindices2<-merge(allindices2,sitedataclyde,by="site",all=T)
allindices2$Clydedistkm<-allindices2$newClydedist0/1000
allindices2$Clydedist2km<-allindices2$newClydedist2/1000

##rescale variables
allindices2<-as.data.frame(allindices2)
allindices2  <- transform(allindices2,logKelvindistr=scale(logKelvindist,center=FALSE),Clydedistkmr=scale(Clydedistkm,center=FALSE),julianr=scale(julian,center=FALSE),timehrr=scale(timehr,center=FALSE),Elevationr=scale(Elevation,center=FALSE),Tempr=scale(Temp,center=FALSE),Rel.Humidityr=scale(Rel.Humidity,center=FALSE),dist_streetlightr=scale(dist_streetlight,center=FALSE),prop_woodlandr=scale(prop_woodland,center=FALSE),prop_wetlandr=scale(prop_wetland,center=FALSE),prop_urbanr=scale(prop_urban,center=FALSE),prop_grasslandr=scale(prop_grassland,center=FALSE),prop_arabler=scale(prop_arable,center=FALSE),dist_allotmentr=scale(dist_allotment,center=FALSE),dist_cemeteryr=scale(dist_cemetery,center=FALSE),dist_parksr=scale(dist_parks,center=FALSE),dist_motorwayr=scale(dist_motorway,center=FALSE),dist_majorroadr=scale(dist_majorroad,center=FALSE),dist_minorroadr=scale(dist_minorroad,center=FALSE),dist_canalr=scale(dist_canal,center=FALSE),dist_laker=scale(dist_lake,center=FALSE),dist_riverr=scale(dist_river,center=FALSE),streetlight_densityr=scale(streetlight_density,center=FALSE))


allindices3<-allindices2[complete.cases(allindices2), ]






## create variable for time from sunrise


#try making data frame with full list of dates, 1 coordinate for all sites, and add to full data frame - subset those which have times before sunrise and adjust the date to correspond with the previous day so they have the previous day's sunrise time

#data frame of list of all dates
datetimes<-allindices3$datetime
datetimes<-dmy_hm(datetimes)
min(datetimes)
max(datetimes)
dates<-seq.Date(from = as.Date("2023-03-10 UTC"), to = as.Date(max(datetimes)), by = 'days') 
str(dates)
datesp<-as.POSIXct(dates)

#spatial points object of average lat long
mean(allindices3$lon)
mean(allindices3$lat)
coord <- matrix(c(mean(allindices3$lon), 
mean(allindices3$lat)), nrow=1)
  Coord <- sp::SpatialPoints(coord, proj4string=sp::CRS("+proj=longlat +datum=WGS84"))
  
  #calculate sunrise time for each day
st3<-sunriset(
coord,
datesp,
direction = "sunrise",
POSIXct.out=TRUE
)

#separate first 16 rows (before daylight savings) and rows 17-173 (british summer time)
st31<-st3[1:16,]
st32<-st3[17:173,]
st32$time<-st32$time-3600

#entire data set in GMT
st4<-rbind(st31,st32)
st4$date<-as.Date(st4$time)
st4$sunrise<-st4$time
st4<-subset(st4,select=c(sunrise,date))

allindices3$date<-as.Date(allindices3$date,"%d/%m/%Y")

#166025
allindices4<-merge(allindices3,st4,by="date")
allindices4$datetime<-dmy_hm(allindices4$datetime)
allindices4$diff<-allindices4$datetime-allindices4$sunrise
allindices4$diffn<-as.numeric(allindices4$diff)

prevday<-filter(allindices4,diffn<=0)
sameday<-filter(allindices4,diffn>0)

prevday$date<-prevday$date-1
prevday<-subset(prevday,select=-c(sunrise))
prevday<-merge(prevday,st4,by="date")

allindices5<-rbind(prevday,sameday)
allindices5$timefromsunrise<-allindices5$datetime-allindices5$sunrise
allindices5$minfromsunrise<-as.numeric(allindices5$timefromsunrise)/60
allindices5$hrfromsunrise<-as.numeric(allindices5$minfromsunrise)/60
hist(allindices5$hrfromsunrise)

p <- 24
indicesdf5rs  <- transform(allindices5,minfromsunriser=scale(minfromsunrise,center=FALSE),hrfromsunriser=scale(hrfromsunrise,center=FALSE))
max(indicesdf5rs$minfromsunrise)

#days getting longer by ~10 minutes each day during study period
#try periodicity of both 24 hours and 24.167 hours (1450 minutes)

#scaling of minfromsunrise divides by 828.5704
#scaling of hrfromsunrise divides by 13.8095
#scaling of timehr divides by 13.66089

#24 hours, rescaled by timehr
prs<-24/13.66089

#24 hours, rescaled by hrfromsunrise 
pshrs<-24/13.8095
#24.167 hours, rescaled by hrfromsunrise
pshrso<-24.167/13.8095


indicesdf5rs<- indicesdf5rs %>% 
  group_by(site,date) %>% 
  arrange(hrfromsunriser, .by_group = TRUE) %>% 
  mutate(time_ordinal = row_number()) 


indicesdf5rs$time_ordinalfac<-as.factor(indicesdf5rs$time_ordinal)


## model temp/humidity ~ time + day + (1|site) to get random effect term for each site to use as explanatory variable in big models

##temperature
tempmod1 <- lmer(Temp ~ hrfromsunrise + julian + (1|site), data = indicesdf5rs)
summary(tempmod1)
hist(resid(tempmod1))
plot(tempmod1)
ranef(tempmod1)
sitetemp <- (ranef(tempmod1))$site
sitetemp$sitetemp <- sitetemp$`(Intercept)`
sitetemp$site <- rownames(sitetemp)
sitetempdf <- subset(sitetemp, select = c("site", "sitetemp"))
sitetempdf$site <- as.factor(sitetempdf$site)

hummod1 <- lmer(Rel.Humidity ~ hrfromsunrise + julian + (1|site), data = indicesdf5rs)
summary(hummod1)
hist(resid(hummod1))
plot(hummod1)
ranef(hummod1)
sitehum <- (ranef(hummod1))$site
sitehum$sitehum <- sitehum$`(Intercept)`
sitehum$site <- rownames(sitehum)
sitehumdf <- subset(sitehum, select = c("site", "sitehum"))
sitehumdf$site <- as.factor(sitehumdf$site)

siteenv <- merge(sitetempdf, sitehumdf, by = "site")



indicesdf5rs <- merge(indicesdf5rs, siteenv, by = "site", all = TRUE)


indicesdf5rs  <- transform(indicesdf5rs,sitetempr=scale(sitetemp,center=FALSE),sitehumr=scale(sitehum,center=FALSE))

days <- wday(indicesdf5rs$datetime)
#dates are one day forward - day of week 3 is actually day of week 2
days2 <- days-1


days2[days2 == '0'] <- '7'
days2 <- as.numeric(days2)
daysdf <- as.data.frame(days2)

daysdf$dayofweek <- case_when(
  days2 < 6 ~ "weekday",
  days2 == 6 ~ "weekend"
)

dow <- subset(daysdf, select = c("dayofweek"))

indicesdf5rs$dayofweek <- dow$dayofweek
indicesdf5rs$dayofweek <- as.factor(indicesdf5rs$dayofweek)

indicesdf5rs$sitedate <- paste(indicesdf5rs$site, indicesdf5rs$date, sep = "-")
indicesdf5rs$sitedate <- as.factor(indicesdf5rs$sitedate)


#data set for gams - remove unnecessary variables
indicesdf5rssub <- subset(indicesdf5rs, select = c(BIO, hrfromsunrise, hrfromsunriser, julian, julianr, prop_urban, prop_urbanr, dist_majorroad, dist_majorroadr, streetlight_density, streetlight_densityr, logKelvindist, logKelvindistr, Clydedistkm, Clydedistkmr, dayofweek, datetime, site, sitetempr))
#write.csv(indicesdf5rssub,"kelvin_corridor_acoustic/outputs/full-data/finaldata.csv")

#write.csv(siteenv,"kelvin_corridor_acoustic/outputs/full-data/tempresid.csv")


```







## road distance ~ noise level correlation

```{r}
noise <- read.csv("data/environmental_variables/Noise values sites Glasgow.csv", header = TRUE, sep = ",")
noise$site <- as.factor(noise$site)
sitedata <- read.csv("data/all_site_data.csv", header = TRUE, sep = ",")
sitedata$site <- as.factor(sitedata$site)

noise2 <- subset(noise, select = c(site, mean_noise))
sitedata2 <- subset(sitedata, select = c(site, dist_majorroad))

sitenoise <- merge(noise2, sitedata2, by = "site")

hist(sitenoise$dist_majorroad)
cor.test(sitenoise$mean_noise, sitenoise$dist_majorroad, method = "spearman")
```









# correlation matrix of explanatory variables

```{r}
ex <- read.csv("data/site_data_tidy.csv", sep = ",", header = TRUE)
```


### subset to relevant variables

Kelvindist, prop_urban, dist_majorroad, streetlight_density, Clydedist


```{r}
ex2 <- subset(ex, select = c("Kelvindist", "prop_urban", "dist_majorroad", "streetlight_density", "Clydedist"))

ex2 <- rename(ex2, dist_riparian = "Kelvindist", upstream_dist = "Clydedist")
```


correlation matrix

```{r}
mat <- round(cor(ex2),2)
```


### heatmap

```{r}
library(reshape2)
mat2 <- melt(mat)
head(mat2)

ggplot(data = mat2, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile()
```


tidy

```{r}
# Get lower triangle of the correlation matrix
  get_lower_tri<-function(mat){
    mat[upper.tri(mat)] <- NA
    return(mat)
  }
  # Get upper triangle of the correlation matrix
  get_upper_tri <- function(mat){
    mat[lower.tri(mat)]<- NA
    return(mat)
  }
  
  upper_tri <- get_upper_tri(mat)
upper_tri




# Melt the correlation matrix
mat3 <- melt(upper_tri, na.rm = TRUE)


# Heatmap
ggplot(data = mat3, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_classic()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
```



```{r}
reorder_mat <- function(mat){
# Use correlation between variables as distance
dd <- as.dist((1-mat)/2)
hc <- hclust(dd)
mat <-mat[hc$order, hc$order]
}


# Reorder the correlation matrix
mat <- reorder_mat(mat)
upper_tri <- get_upper_tri(mat)
# Melt the correlation matrix
mat4 <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(mat4, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
# Print the heatmap
print(ggheatmap)


correlationmatrix <- ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))


ggsave(filename = "kelvin_corridor_acoustic/outputs/figures/correlation_matrix.png",
       plot = correlationmatrix,
       device = "png",
       width = 18,
       height = 12,
       units = "cm",
       dpi = 300,
       bg = "white")
```




# maps of environmental variables


### read in data for this

```{r}
finaldata<-fread("outputs/full-data/finaldata.csv", header=T, sep=",")
finaldata$dayofweek <- as.factor(finaldata$dayofweek)
finaldata$site <- as.factor(finaldata$site)
str(finaldata)

#scaling parameters
#hrfromsunrise - divide by 13.8095
hrfromsunrisescale <- 13.8095
#julian - divide by 158.1123
julianscale <- 158.1123
#logKelvindist - divide by 5.046928
logKelvindistscale <- 5.046928
#Clydedistkm - divide by 8.172523
Clydedistkmscale <- 8.172523
#prop_urban - divide by 0.6215798
prop_urbanscale <- 0.6215798
#dist_majorroad - divide by 409.4769
dist_majorroadscale <- 409.4769
#streetlight_density - divide by 23.69028
streetlight_densityscale <- 23.69028
#sitetemp - divide by 0.4687951
sitetemp_scale <- 0.4687951

#read in point data from GIS to use for spatial predictions
pointdata<-fread("data/environmental_variables/buffer_point_data.csv", header=T, sep=",")
#format variables correctly - rescaling etc
pointdata$logKelvindist <- log(pointdata$kelvindist)
pointdata$Clydedistkm <- pointdata$clydedist / 1000
pointdata$logKelvindistr <- pointdata$logKelvindist / logKelvindistscale
pointdata$Clydedistkmr <- pointdata$Clydedistkm / Clydedistkmscale
pointdata$prop_urbanr <- pointdata$urbanprop / prop_urbanscale
pointdata$dist_majorroadr <- pointdata$roaddist / dist_majorroadscale
pointdata$streetlight_densityr <- pointdata$light_dens / streetlight_densityscale

#make base data frame for predictions
basedf <- subset(pointdata, select = c(logKelvindistr, Clydedistkmr, prop_urbanr, dist_majorroadr, streetlight_densityr, X, Y, site))

#add temp resid
tempresid <- read.csv("kelvin_corridor_acoustic/outputs/full-data/tempresid.csv", header=TRUE, sep=",")
tempresid$site <- as.factor(tempresid$site)
tempresid <- subset(tempresid, select = c(site, sitetemp))

basedf <- merge(basedf, tempresid, by = "site")
basedf$sitetempr <- basedf$sitetemp / sitetemp_scale

## read in Kelvin shapefile
kelvin <- st_read("data/environmental_variables/shapefiles/Kelvin.shp")

#clyde
clyde <- st_read("data/environmental_variables/shapefiles/Clyde.shp")

#major roads shapefile
majorroad <- st_read("data/environmental_variables/shapefiles/major_roads.shp")

#streetlights
streetlight <- st_read("data/environmental_variables/shapefiles/streetlightshp.shp")


```


```{r}
basedf$prop_urban <- basedf$prop_urbanr*prop_urbanscale
basedf$streetlight_density <- basedf$streetlight_densityr*streetlight_densityscale
basedf$dist_majorroad <- basedf$dist_majorroadr*dist_majorroadscale


#plot basic features
map1 <-    ggplot() +
  geom_point(data = basedf, aes(x = X, y = Y),color="lightblue") +
  geom_sf(data=kelvin)+
   geom_sf(data=clyde, linewidth=1)+
   geom_sf(data=majorroad, color="red")+
   geom_sf(data=streetlight,size=0.005,color="yellow", alpha=0.1)+
   coord_sf(xlim = c(254530, 259582),  ylim = c(665720, 672872))+
   theme_bw()+
      scale_color_continuous(aes(name="Proportion of Urban Land"))+
      xlab("Longitude")+
      ylab("Latitude")+
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 6))+
  theme(axis.text.y = element_text(size = 6))+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank())



#plot proportion of urban land
map2 <-    ggplot() +
  geom_point(data = basedf, aes(x = X, y = Y, color=prop_urban)) +
  geom_sf(data=kelvin)+
   geom_sf(data=clyde, linewidth=1)+
   coord_sf(xlim = c(254530, 259582),  ylim = c(665720, 672872))+
   theme_bw()+
      scale_color_continuous(aes(name="Proportion of\nUrban Land"))+
      xlab("Longitude")+
      ylab("Latitude")+
  theme(legend.position = c(0.81, 0.53))+
  theme(legend.key.size = unit(0.2, 'cm'))+
  theme(legend.title = element_text(size=6.5, lineheight = 0.8, margin = margin(b=2)))+
  theme(legend.text = element_text(size=6.5))+ theme(legend.background=element_blank())+
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 6))+
  theme(axis.text.y = element_text(size = 6))+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank())





#streetlight density
map3 <-    ggplot() +
  geom_point(data = basedf, aes(x = X, y = Y, color=streetlight_density)) +
  geom_sf(data=kelvin)+
   geom_sf(data=clyde, linewidth=1)+
   coord_sf(xlim = c(254530, 259582),  ylim = c(665720, 672872))+
   theme_bw()+
      scale_color_continuous(aes(name="Streetlight\nDensity"),low="yellow", high="red")+
      xlab("Longitude")+
      ylab("Latitude")+
  theme(legend.position = c(0.79, 0.52))+
  theme(legend.key.size = unit(0.2, 'cm'))+
  theme(legend.title = element_text(size=7, lineheight = 0.8, margin = margin(b=1.5)))+
  theme(legend.text = element_text(size=6.5))+ theme(legend.background=element_blank())+
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 6))+
  theme(axis.text.y = element_text(size = 6))+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank())
map3



#distance from major road
map4 <-    ggplot() +
  geom_point(data = basedf, aes(x = X, y = Y, color=dist_majorroad)) +
  geom_sf(data=kelvin)+
   geom_sf(data=clyde, linewidth=1)+
   coord_sf(xlim = c(254530, 259582),  ylim = c(665720, 672872))+
   theme_bw()+
      scale_color_continuous(aes(name="Distance\nfrom Major\nRoad (m)"),low="white",high="black")+
      xlab("Longitude")+
      ylab("Latitude")+
  theme(legend.position = c(0.81, 0.5))+
  theme(legend.key.size = unit(0.2, 'cm'))+
  theme(legend.title = element_text(size=7, lineheight = 0.8, margin = margin(b=0.2)))+
  theme(legend.text = element_text(size=6.5))+ theme(legend.background=element_blank())+
  theme(axis.text.x = element_text(angle = 45, hjust=1, size = 6))+
  theme(axis.text.y = element_text(size = 6))+
  theme(axis.title.x = element_blank(), axis.title.y = element_blank())






plot4 <- ggpubr::ggarrange(map1,map2,map3,map4, nrow = 2, ncol = 2, labels = c("A","B","C","D"))

plot4
ggsave(filename = "kelvin_corridor_acoustic/outputs/figures/map_environmental_variables.png",
       plot = plot4,
       device = "png",
       width = 12,
       height = 12,
       units = "cm",
       dpi = 300,
       bg = "white")

```


